Combined effects of protein expression variance and correlation on multicomponent systems

### 2020-03-10 Manuscript/Manuscript.Rmd
update

library(knitr)
library(tidyverse)
library(RColorBrewer)
#library(pheatmap)
library(ComplexHeatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
library(cowplot)
library(patchwork)
source("../functions.R")
library(knitcitations)
library(ggplotify)
library(latex2exp)

@@

@@

>>>>

library(knitr)
library(tidyverse)
library(RColorBrewer)
#library(pheatmap)
library(ComplexHeatmap)
library(GGally)
library(VIM)
library(org.Xl.eg.db)
library(GO.db)
library(ggrepel)
library(plotly)
library(cowplot)
library(patchwork)
source("../functions.R")
library(knitcitations)
library(ggplotify)
library(latex2exp)
library(KEGG.db)

@@

```{r group_prots, include=FALSE}
# GO Terms
GO <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(GOterms = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = 'GOALL', 
                     multiVals = 'list'))

allGO <- unnest(GO, cols = GOterms) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(name = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'TERM') %>% as.character(),
         ontology = mapIds(GO.db, 
                     keys = GOterms, 
                     keytype = 'GOID', 
                     column = 'ONTOLOGY') %>% as.character(),
         type = "GO") %>%
  rename(accession = GOterms) %>%
  dplyr::select(type, accession, name, ontology, everything())

# KEGG Pathways
kegg <- plyr::ldply(as.list(org.Xl.egPATH2EG),data.frame) %>% as_tibble() %>%
  rename(kegg_id = ".id", gene_id = "X..i..")

kegg_names <- plyr::ldply(as.list(KEGGPATHNAME2ID),data.frame) %>%
  rename(kegg_name = ".id",
         kegg_id = "X..i..")

kegg <- protImpute %>% dplyr::select(Uniprot) %>% filter(Uniprot %in% keys) %>% 
  mutate(KEGG = mapIds(org.Xl.eg.db, 
                     keys = Uniprot, 
                     keytype = 'UNIPROT', 
                     column = "PATH", 
                     multiVals = 'list'))

allKEGG <- unnest(kegg, cols = KEGG) %>% nest(Uniprot = Uniprot) %>%
  rowwise() %>% mutate(n = length(unlist(Uniprot))) %>% ungroup() %>%
  mutate(name = kegg_names$kegg_name[match(KEGG,kegg_names$kegg_id)],
         type = "KEGG",
         ontology = NA) %>% rename(accession = KEGG) %>%
  dplyr::select(type, accession, name, ontology, Uniprot, n)

allGroups <- rbind(allGO, allKEGG)

# Calculate summary variables
corVector <- function(uniprotList, corMat){
  targets = uniprotList %>% unlist()
  temp = corMat[rownames(corMat) %in% targets,colnames(corMat) %in% targets]
  temp[lower.tri(temp, diag = F)]
}

selectCV <- function(x){
  protVar %>% filter(Uniprot %in% unlist(x)) %>% 
    dplyr::select(cv) %>% as_tibble() %>% as.list()
}

allGroups <- allGroups %>% group_by(accession) %>% 
  dplyr::mutate(cors = corVector(Uniprot, cor) %>% as_tibble() %>% as.list(),
                covs = corVector(Uniprot, cov) %>% as_tibble() %>% as.list(),
                cv = selectCV(Uniprot)) %>% ungroup() %>%
  rowwise() %>% mutate(medCors = mean(cors, na.rm = TRUE),
                       medCV = mean(cv, na.rm = TRUE)) %>%
  ungroup() %>%
filter(n > 3,
       !is.na(medCors),
       !is.na(medCV))

```{r total_var, include=FALSE}
groupCor <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }
}

totVarPair <- function(x){
  A <- as.character(x[1])
  B <- as.character(x[2])
  varA <- as.numeric(protVar[protVar$Uniprot == A, 3])
  varB <- as.numeric(protVar[protVar$Uniprot == B, 3])
  corAB <- cor[A,B]
  tot <- varA^2 + varB^2 + 2*corAB*varA*varB
  return(tot)
}

totVarGroup <- function(x){
  geneList <- unlist(x)

  if (length(geneList) > 1) {
    protCombs <- t(combn(geneList, 2))
    return(sqrt(sum(apply(protCombs, 1, function(x) totVarPair(x)))))
    #return(sqrt(sum(totVarPair(protCombs))))
  } else{
    return(as.numeric(protVar[protVar$Uniprot == geneList[1], 2]))
  }

}

allGroups <- allGroups %>% filter(!is.na(accession)) %>% rowwise() %>%
  mutate(totVar = totVarGroup(Uniprot),
         normVar = totVar / n) %>%
  filter(n > 3) %>% ungroup()
```

@@

```{r figure_2,, echo=FALSE, fig.height=8, fig.width=6, message=FALSE, warning=FALSE, fig.cap=fig_legend_2}
allGroups <- allGroups %>% mutate(interval = cut_number(normVar, n = 10)) %>%
  mutate(interval = ifelse(interval == levels(interval)[1],
                           "low",
                           ifelse(interval == levels(interval)[10],
                                  "high",
                                  "mid")))

ggplot(allGroups, aes(x = normVar)) + geom_histogram(bins = 100, aes(fill = interval)) + theme_minimal() +
  #scale_fill_viridis_d()
  scale_fill_brewer(palette = "PiYG")

ggplot(allGroups, aes(x = medCors, y = medCV, color = interval)) + 
  geom_point() + 
  theme_minimal()

#fig_legend_2 <- c("A) ")
```
